---
title: "p8105_hw2_ml4239.git"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(readxl)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

#Problem 1 NYC transit
##Import and clean &retain some data; 
retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 

```{r}

NYC_transit_data<-read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) 
```
convert entry variable to from character (`rYES` vs `rNO`) to a logical variable

```{r}
ifelse(NYC_transit_data$entry == "YES", TRUE, FALSE) %>% 
  str()
```

##The NYC_transit_data dataset contains the following variables:

`r names(NYC_transit_data)`

* line: tracks that trains run over
* station_name: station names
* station_latitude: the latitude location of the station
* station_longitude: the longitude location of the station
* route 1-11: route the station serves
* entry: logical vector to show if the station has entry (TRUE) or not (FALSE)
* vending: if the station has vending system or not
* entrance_type: the type of entrance (stair, elevator, door, etc.)
* ada: if the station is ADA compliant

##The step that I used to clean the data:
After importing the data, I used  `clean_names` from `janitor package` to clean the names. And then I select some variables from the original dataset, and convert the 'entry' variable from character to logical variable.

##overview about the dataset
The dimension of the resulting dataset is `r dim(NYC_transit_data)` after cleaning. 

These datas are not tidy because the route number and route name variable are not in their own columns. The variables from route 1 to 11 can be put into the same column as "route_name"variable.
