---
title: "p8105_hw2_ml4239"
Author: "Man Luo"
Date: 2018-10-05
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(ggridges)
library(readxl)
knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)
```

#Problem 1 NYC transit
##Import and clean &retain some data; 
retain line, station, name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. 

```{r}

NYC_transit_data<-read_csv("./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>% 
  janitor::clean_names() %>% 
  select(line, station_name, station_latitude, station_longitude, route1:route11, entry, vending, entrance_type, ada) 
```
convert entry variable to from character (`rYES` vs `rNO`) to a logical variable

```{r}
ifelse(NYC_transit_data$entry == "YES", TRUE, FALSE) %>% 
  str()
```

##The NYC_transit_data dataset contains the following variables:

`r names(NYC_transit_data)`

* line: tracks that trains run over
* station_name: station names
* station_latitude: the latitude location of the station
* station_longitude: the longitude location of the station
* route 1-11: route the station serves
* entry: logical vector to show if the station has entry (TRUE) or not (FALSE)
* vending: if the station has vending system or not
* entrance_type: the type of entrance (stair, elevator, door, etc.)
* ada: if the station is ADA compliant

##The steps that I used to clean the data:
After importing the data, I used  `clean_names` from `janitor package` to clean the names. And then I select some variables from the original dataset, and convert the 'entry' variable from character to logical variable.

##Overview about the dataset
The dimension of the resulting dataset is `r dim(NYC_transit_data)` after cleaning. 

These datas are not tidy because the route number and route name variable are not in their own columns. The variables from route 1 to 11 can be put into the same column as "route_name"variable.

##Questions using these data
* How many distinct stations are there? 
```{r}
nrow(distinct(NYC_transit_data, station_name, line ))
```
There are 465 distinct stations which characterized by name and line.

* How many stations are ADA compliant?
```{r}
NYC_transit_data %>% 
filter(ada==TRUE) %>% 
  distinct(station_name,line) %>% 
  nrow()
```

84 stations are ADA compliant. 

* What proportion of station entrances / exits without vending allow entrance?
```{r}
  sum(NYC_transit_data$vending == "NO"& NYC_transit_data$entry == "YES")/sum (NYC_transit_data$vending == "NO")
```

The proportion of station entrances/ exits without vending allow entrance is 0.3770

##Reformat data 
so that route number and route name are distinct variables
```{r}
NYC_transit_tidy<-gather(NYC_transit_data, key = route_number, value = route_name, route1:route11) 
 
```

* How many distinct stations serve the A train?

```{r}
NYC_transit_tidy %>% 
  filter(route_name == "A") %>%
  distinct(station_name, line) %>% 
  nrow() 
```

60 dinstinct stations serve the A train.

* How many of the stations that serve the A train are ADA compliant?
 
```{r}
NYC_transit_tidy %>% 
filter(route_name == "A" & ada == TRUE) %>% 
  distinct(station_name,line) %>% 
  nrow()
```

17 stations that serve the A train are ADA compliant

#Problem 2 Mr. Trash Wheel
##Import and clean data Mr. Trash Wheel
```{r}
Mr.Trash_data<-read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 1, range = "A2:N258") %>% 
  janitor::clean_names() %>% 
  filter(!is.na(dumpster)) %>% 
  mutate(sports_balls = as.integer(ceiling(sports_balls))) 
 
```

##Read and clean precipitation data for 2016 and 2017.
```{r}
Precipitation_2016_data<-read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 5, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate( year = "2016")

Precipitation_2017_data<-read_excel("./data/HealthyHarborWaterWheelTotals2018-7-28.xlsx", sheet = 4, range = "A2:B15") %>% 
  janitor::clean_names() %>% 
  na.omit() %>% 
  mutate( year = "2017")
```

combine datasets and convert month to a character variable 
```{r}
Precipitation_bind<-
  rbind(Precipitation_2016_data,Precipitation_2017_data) %>% 
  mutate(month = month.name[as.integer(.$month)])
```
 

